Increasing the topological gap by an order of magnitude through the geometry for Majorana SNS junctions.

### 2019-02-07 paper/snakemajoranas.tex
make variables x,y dependent

\comment{Majoranas are mostly made in hybrid NS structures.}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of MBSs in hybrid normal-superconductor (NS) structures, where a long strip (or wire) of a semiconductor is interfaced with a superconductor.~\cite{lutchyn_majorana_2010,oreg_helical_2010}
Within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple wire), the semiconductor becomes topological while MBSs appear at the ends of the segment.
Recently, a system with a slight modification has been proposed\cite{pientka2017topological}, where instead of having one superconductor, there are two superconductors, creating a superconductors-normal-superconductor (SNS) junction.
The additional superconductor adds an extra knob to adjust, the superconducting phase difference $\phi$, and this should make it easier to tune the system into the topological phase.
Recently, two groups have realized this system experimentally.~\cite{fornieri_evidence_2018,ren_topological_2018}
In practice, multiple problems arise making even the observation difficult, leaving the manipulation of Majoranas problematic.

\comment{The gap is small because of long trajectories.}
One of the biggest challenges in creating stable Majoranas is the appearance of a soft gap; where the gap in the density of states significantly reduces for states with the momentum directed along the length of the strip.
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor.
Additionally, these long trajectories have long flight times $\tau_f$ and equivalenly small Thouless energies $E_{\textrm{Th}}=\hbar / \tau_f$, resulting in a small gap.

@@

\comment{We consider a two-dimensional system with zigzag.}
We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two-dimensional strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{equation}
\begin{split}
    H = & \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right)\tau_z \\
        & + E_\text{z} \sigma_x
          + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z
          + \Delta \tau_y.
\end{split}
\label{eq:hamiltonian}
\end{equation}
It acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.
Unless noted differently, the Hamiltonian parameters are $\alpha=\SI{20}{\meV \nm}$, $g=26$, $\meff=\SI{0.02}{\electronmass}$, $\mu=\SI{10}{\meV}$, $B=\SI{1}{T}$, $\phi=\pi$, and $\Delta=\SI{1}{\meV}$; and the geometry parameters are $W=\SI{200}{\nm}$, the period of the zigzag $z_x=\SI{1300}{\nm}$, the discretization contant $a=\SI{10}{\nm}$, and the lengths of the superconductors $L_\textrm{SC}=\SI{200}{\nm}$.

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/bandstructures}
\caption{Figure of the bandstuctures corresponding to the system in Fig.~\ref{fig:setup}.
Blue lines correspond to $\phi=0$, $B=0$ and red lines to $\phi=\pi$, $B = \SI{1}{T}$.
The three subplots are for different amplitudes of the zigzag, with (a) $z_y=0$, (b) $z_y=\frac{W}{4}$, and (c) $z_y=\frac{W}{2}$, where $W=\SI{200}{\nm}$ is the junction width.
We observe that once there are no more straight trajectories inside the junction (when $z_y=\frac{W}{2}$) the spectrum becomes most insensitive to the momentum $k_x$.
The parameters values are written on the bottom of Sec.~\eqref{sec:physical_picture}.\label{fig:bandstuctures}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\lambda_\textrm{M}$ and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi$ for (a) is significantly longer and $E_\textrm{gap} = E_\textrm{1} - E_M$ orders of magnitude smaller than for (b), (c), and (d), meaning that the details of the geometry do not matter for the improvements to occur.
The parameters values are written on the bottom of Sec.~\eqref{sec:physical_picture}.\label{fig:wavefunctions}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/phasediagrams}
\caption{Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
We compute the dispersion of the translationally invariant realizations of a straight and saw-toothed zigzag system and find $\min{E(k)}$ to obtain the energy gap.
% Additionally, use a generalized eigenvalue problem to find all phase boundaries at once.
The parameters values are written on the bottom of Sec.~\eqref{sec:physical_picture}, except with $a=\SI{5}{\nm}$ and $L_\textrm{SC}=\SI{800}{\nm}$.
The chemical potential is $\mu=\SI{10}{\meV}$ in (b) and (d) and the superconducting phase is $\phi=\pi$ in (a) and (c).
\label{fig:phasediagrams}}
\end{figure}

>>>>

\comment{Majoranas are mostly made in hybrid NS structures.}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of MBSs in hybrid normal-superconductor (NS) structures, where a long strip (or wire) of a semiconductor is interfaced with a superconductor.~\cite{lutchyn_majorana_2010,oreg_helical_2010}
Within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple wire), the semiconductor becomes topological while MBSs appear at the ends of the segment.
Recently, a system with a slight modification has been proposed\cite{pientka2017topological}, where instead of having one superconductor, there are two superconductors, creating a superconductors-normal-superconductor (SNS) junction.
The additional superconductor adds an extra knob to adjust, the superconducting phase difference $\phi$, and this should make it easier to tune the system into the topological phase.
Recently, two groups~\cite{fornieri_evidence_2018,ren_topological_2018} have realized this system experimentally.
In practice, multiple problems arise making even the observation difficult, rendering the manipulation of Majoranas in the near future, unprobable.

\comment{The gap is small because of long trajectories.}
One of the biggest challenges in creating stable Majoranas is the appearance of a soft gap; where the gap in the density of states significantly reduces for states with the momentum directed along the length of the strip.  % TODO: do we cite here?
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor.
Additionally, these long trajectories have long flight times $\tau_f$ and equivalenly small Thouless energies $E_{\textrm{Th}}=\hbar / \tau_f$, resulting in a small gap.

@@

\comment{We consider a two-dimensional system with zigzag.}
We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two-dimensional strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{equation}
\begin{split}
    H = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z \\
        & + E_\text{z}(x,y) \sigma_x \\
        & + \alpha(x,y) \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z
          + \Delta(x,y) \tau_y.
\end{split}
\label{eq:hamiltonian}
\end{equation}
It acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.

Unless noted differently, the Hamiltonian parameters are $\alpha=\SI{20}{\meV \nm}$, $g=26$, $\meff=\SI{0.02}{\electronmass}$, $\mu=\SI{10}{\meV}$, $B=\SI{1}{T}$, $\phi=\pi$, and $\Delta=\SI{1}{\meV}$; and the geometry parameters are $W=\SI{200}{\nm}$, the period of the zigzag $z_x=\SI{1300}{\nm}$, the discretization contant $a=\SI{10}{\nm}$, and the lengths of the superconductors $L_\textrm{SC}=\SI{200}{\nm}$.

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/bandstructures}
\caption{Figure of the bandstuctures corresponding to the system in Fig.~\ref{fig:setup}.
Blue lines correspond to $\phi=0$, $B=0$ and red lines to $\phi=\pi$, $B = \SI{1}{T}$.
The three subplots are for different amplitudes of the zigzag, with (a) $z_y=0$, (b) $z_y=\frac{W}{4}$, and (c) $z_y=\frac{W}{2}$, where $W=\SI{200}{\nm}$ is the junction width.
We observe that once there are no more straight trajectories inside the junction (when $z_y=\frac{W}{2}$) the spectrum becomes most insensitive to the momentum $k_x$.
The parameters values are written on the bottom of Sec.~\ref{sec:physical_picture}.\label{fig:bandstuctures}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\lambda_\textrm{M}$ and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi$ for (a) is significantly longer and $E_\textrm{gap} = E_\textrm{1} - E_M$ orders of magnitude smaller than for (b), (c), and (d), meaning that the details of the geometry do not matter for the improvements to occur.
The parameters values are written on the bottom of Sec.~\ref{sec:physical_picture}.\label{fig:wavefunctions}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/phasediagrams}
\caption{Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
We compute the dispersion of the translationally invariant realizations of a straight and saw-toothed zigzag system and find $\min{E(k)}$ to obtain the energy gap.
% Additionally, use a generalized eigenvalue problem to find all phase boundaries at once.
The parameters values are written on the bottom of Sec.~\ref{sec:physical_picture}, except with $a=\SI{5}{\nm}$ and $L_\textrm{SC}=\SI{800}{\nm}$.
The chemical potential is $\mu=\SI{10}{\meV}$ in (b) and (d) and the superconducting phase is $\phi=\pi$ in (a) and (c).
\label{fig:phasediagrams}}
\end{figure}
